---
title: "Ventisqueros code"
author: "Andrea"
date: "2023-09-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(myClim)
library(ggplot2)
```

# OTC TMS Oct 2020- Jul 2023 dataset
# Load and manage data
```{r}
setwd("~/Dropbox/Chirripo/data/micromet data/datos TOMST sensors/tomst-sensors")

# Read with metadata. check ?mc_read_data to make sure your files have the same column order they require
ft <- read.table("/home/andrea/Dropbox/Chirripo/data/micromet data/datos TOMST sensors/2023-07-21 datos julio 2023/files_table_2020_2023_linux.csv", sep=",", header = T) #key: spreadsheet including filepaths of repeated sensor datasets with serial numbers in filename
lt <- read.table("/home/andrea/Dropbox/Chirripo/data/micromet data/datos TOMST sensors/2023-07-21 datos julio 2023/localities_table.csv", sep=",", header = T)

tms.a <- mc_read_data(files_table = ft,
                      localities_table = lt,
                      silent = T)
#mc_info_count(tms.a) #Basic information about how many sensors etc. you have. 
#mc_info(tms.a) #returns dataframe with summary per sensor
names(tms.a)
```

## Change from solar time to UTC
```{r}
# This needs to happen before aggregating...but I need to do it before otherwise agg is wrong
tms.b <- mc_prep_solar_tz(tms.a) # calculate solar time. I think it only works for aggregated data
#mc_info(tms.b)
#?mc_prep_meta_locality()
#?mc_prep_solar_tz
```

## Join different datasets from the same sensor
```{r}
tms.all <- mc_join(tms.a)
#mc_info_count(tms.all) # Check it worked
#mc_info(tms.all)
```

## Summarize data on a per day basis
```{r}
tms.day <- mc_agg(tms.all, fun = "mean", period = "day", min_coverage=0.1)
#mc_info(tms.day)
```

## Eyeball data through plotting
```{r}
tms.plot <- mc_filter(tms.all)
p<-mc_plot_line(tms.plot, filename="files.pdf", sensors="TMS_TMSmoisture")
p
```

## Convert raw counts (TDR) into VWC
```{r}
tms.vwc.uni<-mc_calc_vwc(tms.all, soiltype="universal") # Using universal equation
#mc_info(tms.vwc.uni) #returns dataframe with summary per sensor
```

# Summarize 
## Entire period
### Aggregate
```{r}
# Aggregate: return one VWC value per sensor for the entire study period
tms.logger.vwc<-mc_filter(tms.vwc.uni, sensor="VWC_moisture")
tms.logger <- mc_agg(tms.logger.vwc, fun = "mean", period = "all", min_coverage = 0.95)
#mc_info(tms.logger)
r <- mc_reshape_long(tms.logger)
#View(r) # Check that it worked
write.csv(r, "/home/andrea/Dropbox/Chirripo/data/micromet data/datos TOMST sensors/2023-07-21 datos julio 2023/VWC_per_logger.csv", 
          row.names = FALSE) 
str(r)

r$treat1<-regexpr("con", r$locality_id) ## Create a new column indicating treatment.If r$locality_id contains "con" return "-1. 
View(r)
r$Treatment <- ifelse(r$treat1==-1, "war","con") # Create a new column that says "war" when there's a -1 and con anytime else
r$block<-substr(r$locality_id, 1, 1) # Create a new column with block
```
### Plot total VWC
```{r}
vwcAll <- ggplot(r, aes(x = block, y = value , fill = Treatment)) +
  geom_boxplot(alpha=0.7) +
  xlab("Block") + ylab("Volumetric soil moisture (%)") +  # label the x & y axes
  ggtitle("Average soil moisture - All seasons\n 2020-2023") + # add a title
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Arial"),
        text = element_text(size = 12, family = "Arial"), 
        panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank(),
        axis.text.x=element_text(size = 12)) +
  scale_fill_manual(values = c("blue", "red"))
vwcAll

ggsave(filename="PlotMeanLoggerVWC.tiff", plot = vwcAll, device = "tiff",
       path = NULL, scale = 1, width = 8, height = 10, units = "cm", 
       limitsize = TRUE, bg = NULL,
)

model<-aov(r$value~r$treat*r$block) # Statistics
summary(model)
plot(model)
```

## Dry season
### Aggregate 
```{r}
# Aggregate: return one VWC value per sensor for the dry season only 
tms.logger.dry <- mc_agg(tms.logger.vwc, fun = "mean",
                  percentiles = 95, period = "custom", min_coverage = 0.95, 
                  custom_start="12-01", custom_end = "04-01")
rdry <- mc_reshape_long(tms.logger.dry)
#View(rdry) # Check that it worked
write.csv(rdry, "/home/andrea/Dropbox/Chirripo/data/micromet data/datos TOMST sensors/2023-07-21 datos julio 2023/VWCperLoggerDrySeason.csv", 
          row.names = FALSE) 

# Create a new column indicating treatment
rdry$treat1<-regexpr("con", rdry$locality_id) # if r$locality_id contains "con" return "-1. 
#View(rdry)
rdry$Treatment <- ifelse(rdry$treat1==-1, "war","con") # Create a new column that says "war" when there's a -1 and con anytime else

rdry$block<-substr(rdry$locality_id, 1, 1) # Create a new column with block
rdry$Year<-format(rdry$time_to, format="%Y") # Create a new column with year
```

### Plot dry season VWC
```{r}
vwcDry <- ggplot(rdry, aes(x = block, y = value , fill = Treatment)) +
  geom_boxplot(alpha=0.7) +
  xlab("Block") + ylab("Volumetric soil moisture (%)") +  # label the x & y axes
  ggtitle("Average soil moisture - Dry season\n 2021-2023") + # add a title
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Arial"),
        text = element_text(size = 12, family = "Arial"), 
        panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank(),
        axis.text.x=element_text(size = 12)) +
  scale_fill_manual(values = c("blue", "red"))
vwcDry

ggsave(filename="PlotMeanLoggerVWCDry.tiff", plot = vwcDry, device = "tiff",
       path = NULL, scale = 1, width = 12, height = 10, units = "cm", 
       limitsize = TRUE, bg = NULL,
)

model<-aov(rdry$value~rdry$treat*rdry$block) # Statistics
summary(model)
plot(model)
```

## Wet season 
### Aggregate
```{r}
# Aggregate: return one VWC value per sensor for the wet season only 
tms.logger.wet <- mc_agg(tms.logger.vwc, fun = "mean",
                         percentiles = 95, period = "custom", min_coverage = 0.95, 
                         custom_start="04-02", custom_end = "11-30")
rwet <- mc_reshape_long(tms.logger.wet)
#View(rwet) # Check that it worked
write.csv(rwet, "/home/andrea/Dropbox/Chirripo/data/micromet data/datos TOMST sensors/2023-07-21 datos julio 2023/VWCperLoggerWetSeason.csv", 
          row.names = FALSE) 

# Create a new column indicating treatment
rwet$treat1<-regexpr("con", rwet$locality_id) # if r$locality_id contains "con" return "-1.
#View(rwet)
rwet$Treatment <- ifelse(rwet$treat1==-1, "war","con") # Create a new column that says "war" when there's a -1 and con anytime else

rwet$block<-substr(rwet$locality_id, 1, 1) # Create a new column with block
rwet$Year<-format(rwet$time_to, format="%Y") # Create a new column with year
```

### Plot wet season
```{r}
vwcWet <- ggplot(rwet, aes(x = block, y = value , fill = Treatment)) +
  geom_boxplot(alpha=0.7) +
  xlab("Block") + ylab("Volumetric soil moisture (%)") +  # label the x & y axes
  ggtitle("Average soil moisture - Wet season\n 2020-2023") + # add a title
  theme_bw() +
  theme(plot.title = element_text(size = 14, family = "Arial"),
        text = element_text(size = 12, family = "Arial"), 
        panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank(),
        axis.text.x=element_text(size = 12)) +
  scale_fill_manual(values = c("blue", "red"))
vwcWet

ggsave(filename="PlotMeanLoggerVWCWet.tiff", plot = vwcWet, device = "tiff",
       path = NULL, scale = 1, width = 12, height = 10, units = "cm", 
       limitsize = TRUE, bg = NULL,
)
getwd()

# Statistics
model<-aov(rwet$value~rwet$treat*rwet$block)
summary(model)
plot(model)

```


## Extract data from certain dates
```{r echo = FALSE, eval = FALSE}
# Extract data from Feb 2 and 3, 2022 only to compare with respiration data
start <- as.POSIXct("2022-02-03")
end <- as.POSIXct("2022-02-03")
tms.feb3 <- mc_prep_crop(tms.v, start, end)
#mc_info(tms.feb3)

write.csv(t,"/home/andrea/Dropbox/Chirripo/data/micromet data/datos TOMST sensors/2023-07-21 datos julio 2023/raw_moisture_feb_2_2022.csv", 
          row.names = FALSE)

# Aggregate data for Feb 2 and 3 only to compare with gravimetric soil moisture
tms.all.feb3<-mc_filter(tms.feb3, sensors=c("TMS_TMSmoisture"))
#mc_info(tms.feb3)
tms.all.feb3<- mc_agg(tms.all.feb3, fun = "mean", period = "all", 
                       min_coverage = 0.95)
t <- mc_reshape_long(tms.feb3)
#View(t)
write.csv(r,"/home/andrea/Dropbox/Chirripo/data/micromet data/datos TOMST sensors/2023-07-21 datos julio 2023/raw_moisture_feb_2_3_2022.csv", 
          row.names = FALSE)
```

